---
title: "PA1_template"
author: "Gabe Rezavi"
date: "7/25/2019"
output: html_document
---

This is my second attempt for our RMarkdown Project 1 document. Let's load the data and libraries and check it out.

```{r}
library(ggplot2)
library(dplyr)
activity <- read.csv(file="activity.csv", header=TRUE, sep=",")
str(activity)
head(activity,3)
```
What is mean total number of steps taken per day? Visualization (histogram)? 

```{r}
Tot.Steps <- aggregate(steps ~ date, activity, FUN=sum)
head(Tot.Steps,3)
hist(Tot.Steps$steps, col="blue", xlab = "Number of Steps", ylab = "Steps", main = "Total Steps taken by Day")
```
Mean and Median of the total number of steps taken per day?

```{r}
Mean.Steps <- mean(Tot.Steps$steps)
Mean.Steps
Median.Steps <- median(Tot.Steps$steps)
Median.Steps
```

What is the average daily activity pattern?

```{r}
Intervals <- aggregate(steps ~ interval, activity, FUN=sum)
head(Intervals,3)
plot(Intervals$interval, Intervals$steps, type = "l", lwd = 2, xlab = "Interval", ylab = "Total Steps", main = "The Average Daily Activity Pattern") 
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
max.Steps <- Intervals[which.max(Intervals$steps),1]
max.Intervals <- max(Intervals)
max.Steps
max.Intervals
```
The maximum numbers of steps `r max.Intervals` were taken on the `r max.Steps` interval, as illustrated here using the inline code.

Now, I will try to impute missing values starting with first calculating the total number of missing values (NAs) in the dataset.

```{r}
table(is.na(activity))
```

Let's impute the missing values with an appropriate strategy and recreate the data set as max.Steps2.

```{r}
Mean.Intervals <- aggregate(steps ~ interval, activity, FUN=mean)
New.Intervals <- merge(x=activity, y=Mean.Intervals, by="interval")

New.Intervals$steps <- ifelse(is.na(New.Intervals$steps.x), New.Intervals$steps.y, New.Intervals$steps.x)
head(New.Intervals)

New.Intervals <- select(New.Intervals, steps, date, interval)
head(New.Intervals)

Tot.Steps2 <- aggregate(steps ~ date, New.Intervals, FUN=sum)
head(Tot.Steps2,3)
```

Before and after visualization via simple histograms.

```{r}
par(mfrow=c(1,2))

hist(Tot.Steps2$steps, col="yellow", xlab = "Steps", ylab = "Frequency", ylim = c(0,35), main = "Total Number Of Steps Taken Each day (w/ Imputation)", cex.main = 0.7)

hist(Tot.Steps$steps, col="red", xlab = "Steps", ylab = "Frequency", ylim = c(0,35), main = "Total Number Of Steps Taken Each day (w/out Imputation)", cex.main = 0.7)
```
Now, let's calculate how much the mean and the media values have changed as a result of these imputations.

```{r}
Mean.Steps2 <- mean(Tot.Steps2$steps)
Median.Steps2 <- median(Tot.Steps2$steps)
Diff.in.Mean <- Mean.Steps - Mean.Steps2
Diff.in.Median <- Median.Steps - Median.Steps2
Diff.in.Mean
Diff.in.Median
```

The difference in mean is `r Diff.in.Mean` while the difference in the median is `r Diff.in.Median`.

Now, let's check out the difference in activities between the weekdays and the weekend.

```{r}
library(chron)
table(is.weekend(New.Intervals$date))
New.Intervals$dayofweek <- ifelse(is.weekend(New.Intervals$date), "weekend", "weekday")
head(New.Intervals)

Mean.Itervals2 <- aggregate(steps ~ interval + dayofweek, New.Intervals, FUN=mean)
head(Mean.Itervals2)

ggplot(Mean.Itervals2, aes(x=interval, y=steps)) + geom_line(color="blue", size=1) + facet_wrap(~dayofweek, nrow=2) + labs(x="\nInterval", y="\nNumber of steps")
```
Thank you!

